Case study validations of automatic bullet matching

### 2018-11-19 index.Rmd
rates to percentages

Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength. Maximized cross-correlation is used to pair-wise align signatures \citet{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.

@@

Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting. Error rates reported by random forest algorithms are based on these internal test sets. However, there is some discussion on whether or in which situations these error rates are biased \citep{rfOver2, rfOver}. 
Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability). For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data.Good performance (in terms of low error rates) on external test data validates a ML algorithm. External data also allows an assessment of the algorithm's sensitivity to distributional changes as well as testing the algorithm's robustness by going outside the parameters of the training data.

@@

Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study. The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set. The Phoenix PD set uses Ruger P-95 barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels. The Houston sets use Ruger LCP barrels. These barrels are first rifled traditionally, i.e. similar to the Ruger P-85, but are treated to a secondary round of heating after \svp{rifling}. This second round of heat exposure has the potential to introduce some subclass characteristics, i.e. anomalies that are shared between different barrels manufactured in this manner. This should make the automatic matching harder, and in particular, \svp{is expected to complicate} the classification as different-source land engraved areas/bullets. \svp{In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.}

@@

h44nest <- h44nest %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  }),
  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
    max(scores)
  })
)
h44nest <- h44nest %>%
  mutate(
    barrel2 = factor(barrel2, levels = c(1:10, "Unknown")),
    bullet1 = factor(bullet1, levels = rev(c(
      "K", "O", "L", "P",
      "J",
      "H", "I", "Y", "G",
      "E", "U", "X", "F",
      "T", "S"
    ))),
    bullet2 = factor(bullet2, levels = c(1:2, rev(levels(bullet1))))
  )
h44 <- h44nest %>%
  ggplot(aes(
    y = bullet1, x = bullet2,
    fill = sam_rf
  )) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel2, space = "free", scales = "free") +
  ylab("Questioned bullets") +
  xlab("Testfires from barrels 1 to 10 (left), questioned bullets (right)") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.35, limits = c(0, 1)
  ) +
  scale_colour_manual("same source", values = "darkorange") +

@@

\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).} \svp{As the land-to-land scores are not perfectly separated, we can use ROC curves to evaluate their performance.} \svp{The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a). A perfect classifier would have 100\% true positives and 0\% false positives, e.g. the ROC curve would be a right angle at (0, 1). Classifiers with better performance will be closer to this corner of the plot. A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}

```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons of the random forest score."}
library(plotROC)
library(pROC)
rocpic <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>%
  mutate(
    measure = factor(measure, labels = c("CC", "CMS", "RF")),
    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))
  ) %>%
  ggplot(aes(linetype = measure, colour = studyg)) +
  geom_roc(aes(m = score, d = as.numeric(sameland)), 
           labels = FALSE, pointsize = 0, alpha = 0.8, size = .75) +
  theme_bw() +
  # coord_fixed() + 
  facet_wrap(~studyg, nrow=3) +
  xlab("Fraction False Positives") +
  ylab("Fraction True Positives") +
  scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
                                         "#74a9cf", "#2b8cbe", "#045a8d"),
                      guide=FALSE) +
  theme(legend.position= c(1, 0), legend.justification = c(1, 0), legend.direction = "vertical",
        legend.key.width = unit(1, "cm")) +
  guides(linetype = guide_legend(override.aes = list(shape = NA, size = .5))) + 
  scale_linetype_manual("Measure", values=c(1,3,2)) 
```

@@

```{r roc-auc, echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).   As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study. Based on AUC, the overall performance on all sets is very good to excellent for cross-correlation and the random forest score. At the land level there is no significant difference in prediction power between RF score and cross-correlation."}
  rocpic
  aucpic
```

@@

\hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) and their thresholds based on the ROC curves of \autoref{fig:roc-auc}. EERs are error rates when sensitivity and specificity of a test are equal, i.e. we see the same error rates of missing an identification for a land-to-land comparison  and wrong identifications, when it is determined that two land engraved areas come from the same source when they are, in fact, known to be from a different source. Equal error rates based on CMS of two lands are significantly higher than equal error rates based on cross-correlation or the RF score. At best, RF score and cross-correlation have an equal error rate of around 5\% for land-to-land comparisons. 
Interestingly, for consecutively matching striae, an equal error rate is reached for just two matching striae, which from a practical point of view is not sustainable. A  CMS of  6 or above is an established cutoff used in identifications. This leads to 21 wrong identifications (out of 49286; for an error rate of 4 in 10,000) but only identifies 139 out of  715 known same-source lands, for an error rate of 80.6\%. }
```{r eval=FALSE}
xtabs(~sameland+I(round(cms)>=6), data = all_lands)
```

```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error Rates and thresholds for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies. CMS has an EER threshold of 2 for all studies. CMS is signifcantly worse than both Cross-correlation and the RF score. At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.\\svp{I'm uncomfortable using the same scale for EER threshold given that CMS has a different scale; leaving CMS out also seems odd though}", out.width = '.8\\textwidth'}
aucs %>% gather(eer_thresh, value, eer, threshold) %>%
  filter(!(measure=="CMS" & eer_thresh=="threshold")) %>%
  mutate(
    eer_thresh = factor(eer_thresh, labels=c("EER", "Threshold for EER"))
  ) %>%
  ggplot(aes(y = studyg, x = value, colour=measure, shape = measure)) + 
  geom_point(size = 4) +theme_bw() +
#  xlim(c(0, .35)) +
  xlab("Equal Error Rates (left) and EER thresholds (right)") +
  ylab("") +
  scale_color_brewer(palette = "Set2") +
  scale_shape_discrete(solid = F) + 
  theme(legend.position = "bottom") +
  facet_grid(.~eer_thresh, scales="free") +
  xlim(c(0, NA))
```

>>>>

Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength. Maximized cross-correlation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.

@@

Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting. Errors reported by random forest algorithms are based on these internal test sets. However, there is some discussion on whether or in which situations these errors are biased \citep{rfOver2, rfOver}. 
Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability). For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data. Good performance (in terms of wrong and missed identifications) on external test data validates a ML algorithm. External data also allows an assessment of the algorithm's sensitivity to distributional changes as well as testing the algorithm's robustness by going outside the parameters of the training data.

@@

Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study. The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set. The Phoenix PD set uses Ruger P-95 barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels. The Houston sets use Ruger LCP barrels. These barrels are first rifled traditionally, i.e. similar to the Ruger P-85, but are treated to a secondary round of heating after rifling. This second round of heat exposure has the potential to introduce some subclass characteristics, i.e. anomalies that are shared between different barrels manufactured in this manner. This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets. In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.

@@

h44nest <- h44nest %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  }),
  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
    max(scores)
  })
)
h44nest <- h44nest %>%
  mutate(
    barrel2 = factor(barrel2, levels = c(1:10, "Unknown")),
    bullet1 = factor(bullet1, levels = rev(c(
      "K", "O", "L", "P",
      "J",
      "H", "I", "Y", "G",
      "E", "U", "X", "F",
      "T", "S"
    ))),
    bullet2 = factor(bullet2, levels = c(1:2, rev(levels(bullet1))))
  )
h44 <- h44nest %>%
  ggplot(aes(
    y = bullet1, x = bullet2,
    fill = sam_rf
  )) +
  geom_tile(size = 1) +
  facet_grid(. ~ barrel2, space = "free", scales = "free") +
  ylab("Questioned bullets") +
  xlab("Test fires from barrels 1 to 10 (left), questioned bullets (right)") +
  scale_fill_gradient2("RF score ",
    low = "darkgrey",
    high = "darkorange", midpoint = 0.35, limits = c(0, 1)
  ) +
  scale_colour_manual("same source", values = "darkorange") +

@@

When evaluating classifier performance, it is common to use a receiver operating characteristic (ROC) curve, which plots the percent of wrong identifications against the percent correct identifications for each possible value of the cutoff between the two classes (in this case, known match and known non-match). Land-to-land scores are not perfectly separated, we can therefore use ROC curves to distinguish between the performance of the different methods and studies. The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a). A perfect classifier would have 100\% correct identifications and 0\% wrong identifications, e.g. the ROC curve would be a right angle at (0, 100). Classifiers with better performance will be closer to this corner of the plot. A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (100,100).

```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons of the random forest score."}
library(plotROC)
library(pROC)
rocpic <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>%
  mutate(
    measure = factor(measure, labels = c("CC", "CMS", "RF")),
    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))
  ) %>%
  ggplot(aes(linetype = measure, colour = studyg)) +
  geom_roc(aes(m = score, d = as.numeric(sameland)), 
           labels = FALSE, pointsize = 0, alpha = 0.8, size = .75) +
  theme_bw() +
  # coord_fixed() + 
  facet_wrap(~studyg, nrow=3) +
  xlab("Percent Wrong Identifications") +
  ylab("Percent Correct Identifications") +
  scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
                                         "#74a9cf", "#2b8cbe", "#045a8d"),
                      guide=FALSE) +
  theme(legend.position= c(1, 0), legend.justification = c(1, 0), legend.direction = "vertical",
        legend.key.width = unit(1, "cm")) +
  guides(linetype = guide_legend(override.aes = list(shape = NA, size = .5))) + 
  scale_linetype_manual("Measure", values=c(1,3,2)) +
  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25\\%", "50\\%", "75\\%", "100\\%")) +
  scale_y_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25\\%", "50\\%", "75\\%", "100\\%"))

@@

```{r roc-auc, echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).   As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study. Based on AUC, the overall performance on all sets is very good to excellent for cross-correlation and the random forest score. At the land level there is no significant difference in prediction power between RF score and cross-correlation.", cache=FALSE}
  rocpic
  aucpic
```

@@

\autoref{fig:eer} shows an overview of Equal Error and their thresholds based on the ROC curves of \autoref{fig:roc-auc}. Equal Errors are errors  when sensitivity and specificity of a test are equal, i.e. we see the same percentage of missed and wrong identifications for a land-to-land comparison. Equal errors based on CMS of two lands are significantly higher than equal errors based on cross-correlation or the RF score. At best, RF score and cross-correlation have an equal error of around 5\% for land-to-land comparisons. 
Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable. A  CMS of  6 or above is an established cutoff used in identifications. This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of  715 known same-source lands, for an error of 80.6\%. 
```{r eval=FALSE}
xtabs(~sameland+I(round(cms)>=6), data = all_lands)
```

```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies. CMS has an equal error threshold of 2 for all studies. CMS is signifcantly worse than both Cross-correlation and the RF score. At best, Cross-correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
aucs %>% 
  mutate(
    eer = 100*eer
  ) %>%
  gather(eer_thresh, value, eer, threshold) %>%
  filter(!(measure=="CMS" & eer_thresh=="threshold")) %>%
  mutate(
    eer_thresh = factor(eer_thresh, labels=c("Equal Error (in Percent)", "Threshold for Equal Error"))
  ) %>%
  ggplot(aes(y = studyg, x = value, colour=measure, shape = measure)) + 
  geom_point(size = 4) +theme_bw() +
  xlab("Equal Error (in Percent, left) and corresponding thresholds (right)") +
  ylab("") +
  scale_color_brewer(palette = "Set2") +
  scale_shape_discrete(solid = F) + 
  theme(legend.position = "bottom") +
  facet_grid(.~eer_thresh, scales="free") +
  xlim(c(0, NA))
```
